Numerical Investigation of Viscous Flow Instabilities in Multiphase Heterogeneous Porous Media

### 2018-02-08 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
updates

\medskip
Figures~\ref{fig:HeleShawHeter_VR10} and \ref{fig:HeleShawHeter_VR150} show formation of multi-scale elongations with continuous growth and coalescence of dendritic finger branching in simulations performed with viscosity ratio of 10 and 150, respectively. In both simulations, maximum fingers width of approximately 0.44-0.56 ({\it VR}=10) and 0.24-0.44 cm were found, whereas estimated values (based on Eqn.~\ref{eqn:guaneqn}) were 0.25-4.50 ({\it VR}=10) and 0.88-3.00 cm ({\it VR}=150 -- (Fig.~\ref{fig:homoheleshaw_VN10_VN150} c-d, Table~\ref{table:Heleshaw}). Such calculated ranges \citep[\ie theoretical values based on analytic solutions due to][]{mclean_1981} clearly overestimate fingers dimensions as the mathematical formulation (and therefore Eqn.~\ref{eqn:guaneqn}) assumes incompressible flows in homogeneous porous medium, \ie it does not take into account any spatial variability of heterogeneity.

@@

 is perturbed if fluids have different densitites and/or viscositiesleading to an uneven front with elongations at the outside edge of the fluid interface

viscous fingering occurs when the viscosity ratio is greater than unity. As surface tension becomes weak, the interface is stressed and becomes unstable leading to the formation of fingers.

%(2)Summary (1-3 paragraphs) of the main results with the appropriate analysis (\ie how you work has proved the theory);

Starting from validating the development of fingers in a computational domain against experimental setup a series of other computational cases is showed focusing on the details of the flows and the introduction of parameteres by avoiding any upscaling requirements.

>>>>

\medskip
Figures~\ref{fig:HeleShawHeter_VR10} and \ref{fig:HeleShawHeter_VR150} show formation of multi-scale elongations with continuous growth and coalescence of dendritic finger branching in simulations performed with viscosity ratio of 10 and 150, respectively. In both simulations, maximum fingers width of approximately 0.44-0.56 ({\it VR}=10) and 0.24-0.44 cm were found, whereas estimated values (based on Eqn.~\ref{eqn:guaneqn}) were 0.25-4.50 ({\it VR}=10) and 0.88-3.00 cm ({\it VR}=150) -- Fig.~\ref{fig:homoheleshaw_VN10_VN150} c-d, Table~\ref{table:Heleshaw}. Such calculated ranges \citep[\ie theoretical values based on analytic solutions due to][]{mclean_1981} clearly overestimate fingers dimensions as the mathematical formulation (and therefore Eqn.~\ref{eqn:guaneqn}) assumes incompressible flows in homogeneous porous medium, \ie it does not take into account any spatial variability of heterogeneity.

@@

%(2)Summary (1-3 paragraphs) of the main results with the appropriate analysis (\ie how you work has proved the theory);
The numerical formulation was initially validated against laboratory experiments \citep{dawe_2008} to qualitatively assess model functionality to simulate crossflow (\ie preferential flow pathways) in a chequerboard permeability domain. Formation and development of viscous fingers in Hele-Shaw cells (homogeneous domain) were investigated to demonstrate that the onset of interface flow instability occur when VR$\ge$3. Maximum dendritic finger widths obtained from numerical simulations performed with VR=10 and VR=150 largely agree with expected parametric analytic solutions \citep{mclean_1981,guan_2003}. However, when perturbations are imposed into the domain geophysical properties, \ie spatial variability of absolute permeability, numerical solutions partially agree with analytic solutions. This may be due to the domain homogeneity assumption in the analytic solution (Eqns.~\ref {eqn:guaneqn}).

analytic solution seems to overestimate maximum finger widths as theoverestimate   dendritic finger branching start to form as predicted by \citep{mclean_1981}. However
